---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r load, echo=TRUE}
unzip(zipfile = "activity.zip")
data <- read.csv("activity.csv")
sumData <- aggregate(data$steps, by = list(date = data$date), FUN = sum)
avgData <- aggregate(data$steps, by = list(interval = data$interval), FUN = mean, na.rm = TRUE)
```

## What is mean total number of steps taken per day?

###Histogram of the total number of steps taken each day

```{r histNA, echo=TRUE}
hist(sumData$x , xlab = "Quantity of steps per day", main = "Histogram of the total number of steps taken each day")
```

###Mean and median total number of steps taken per day
```{r meanMedianNA, echo=TRUE}
meanData <- mean(sumData$x, na.rm = TRUE)
medianData <- median(sumData$x, na.rm = TRUE)
```

* Mean: **`r format(meanData, scientific = FALSE)`**
* Median: **`r format(medianData, scientific = FALSE)`**

## What is the average daily activity pattern?

```{r averageDaily, echo=TRUE}
plot(avgData$interval, avgData$x, type = "l", xlab = "5-minute interval", ylab = "Average of steps" , main = "Histogram of average steps per day")
```

###5-minute interval with maximum averaged number of steps
```{r MaximumDaily, echo=TRUE}
interval <- avgData[avgData$x == max(avgData$x),"interval"]
```

The 5-minute interval which has the maximum number of steps averaged by day is **`r interval`**.

## Imputing missing values

### Quantity of missing values
```{r qtyNAs, echo=TRUE}
qty <- sum(is.na(data))
```
The amount of missing values in the data is **`r qty`**.

### New dataset with missing data filled in

The strategy to fill the missing data is to use the average of steps of each day of each interval.

```{r replacingNAs, echo=TRUE}
newData <- data
newData$avg <- rep(avgData$x,61)
newData$steps[is.na(newData$steps)] <- newData$avg[is.na(newData$steps)]
```

#### Histogram of the total number of steps taken each day

```{r hist, echo=TRUE}
sumNewData <- aggregate(newData$steps, by = list(date = newData$date), FUN = sum)
hist(sumNewData$x , xlab = "Quantity of steps per day", main = "Histogram of the total number of steps taken each day")
```

#### Mean and median total number of steps taken per day
```{r meanMedian, echo=TRUE}
meanData <- mean(sumNewData$x, na.rm = TRUE)
medianData <- median(sumNewData$x, na.rm = TRUE)
```

* Mean: **`r format(meanData, scientific = FALSE)`**
* Median: **`r format(medianData, scientific = FALSE)`**

The values of mean/median vary a little about the first dataset, because the new dataset is filled in with mean values of each interval. On the other hand, there are more observations available (because there are `r qty` more values available than before) and thus, in the histogram there are more observations for the mean value of steps per day.

## Are there differences in activity patterns between weekdays and weekends?

### New dataset which differentiates weekdays and weekends
```{r weekdays, echo = TRUE}
library(lubridate)
newData$typeDay <- ifelse(wday(newData$date) == 1 | wday(newData$date) == 7,"weekend","weekday")
newData$typeDay <- factor(newData$typeDay)
```


```{r averageNewDaily, echo=TRUE}
library(lattice)
avgNewData <- aggregate(newData$steps, by = list(interval = newData$interval, typeDay = newData$typeDay), FUN = mean, na.rm = TRUE)
xyplot(x ~ interval| typeDay, data = avgNewData, type = "l", layout = c(1,2), xlab = "Interval", ylab = "Number of steps")
```

